Lineage Analysis of CRISPR and shRNA Screens

### 2018-10-12 lineages_18Q3.Rmd

## DepMap cell line metadata

@@

## CCLE MAF file

@@

# Add Mutation_Status column
maf_df$Mutation_Status_Deleterious <- ifelse(maf_df$isDeleterious == TRUE, "Mutant", "Other")
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Other", "Mutant")
maf_df$Mutation_Status_DeleteriousMissense <- ifelse(maf_df$isDeleterious == TRUE | maf_df$Variant_Classification == "Missense_Mutation", "Mutant", "Other")
maf_df <- unique(subset(maf_df, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Mutation_Status_Deleterious", "Mutation_Status_Nonsilent", "Mutation_Status_DeleteriousMissense", "Reference_Allele_Length", "Tumor_Seq_Allele1_Length")))

@@

## CCLE copy number file

@@

## CCLE gene expression file

@@

## DepMap dependency probabilities

@@

## Merge CRISPR data and annotations

@@

```{r fig.height=12, fig.width=12}
crispr_pca_data <- merge(ccl_converter[, c("Broad_ID", "group_general_lineage_name")], crispr, by = "Broad_ID")
crispr_lin <- crispr_pca_data[, 2]
crispr_pca <- prcomp(data.matrix(crispr_pca_data[, 3:ncol(crispr_pca_data)]), scale. = TRUE, center = TRUE)
plot(crispr_pca, type = "l")
biplot(crispr_pca, obs.scale = 1, var.scale = 1, groups = crispr_lin, ellipse = TRUE, circle = TRUE)
ggord(crispr_pca, crispr_pca_data$group_general_lineage_name)
```

@@

# Merge mutation annotations
crispr_muts <- merge(crispr_melt, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Deleterious = if_else(is.na(Mutation_Status_Deleterious), "Other", Mutation_Status_Deleterious), Mutation_Status_DeleteriousMissense = if_else(is.na(Mutation_Status_DeleteriousMissense), "Other", Mutation_Status_DeleteriousMissense), Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Other", Mutation_Status_Nonsilent))

# Summarize number of mutant and Other cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Deleterious_Other = sum(Mutation_Status_Deleterious == "Other"),
            N_Deleterious_Mutant = sum(Mutation_Status_Deleterious == "Mutant"),
            N_DeleteriousMissense_Other = sum(Mutation_Status_DeleteriousMissense == "Other"),
            N_DeleteriousMissense_Mutant = sum(Mutation_Status_DeleteriousMissense == "Mutant"),
            N_Nonsilent_Other = sum(Mutation_Status_Nonsilent == "Other"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))

@@

# Add Color columns
crispr_data$Color_Deleterious <- ifelse(crispr_data$Mutation_Status_Deleterious == "Other", "cyan3", "darkorchid")
crispr_data$Color_Deleterious <- factor(crispr_data$Color_Deleterious)
crispr_data$Color_DeleteriousMissense <- ifelse(crispr_data$Mutation_Status_DeleteriousMissense == "Other", "cyan3", "darkorchid")
crispr_data$Color_DeleteriousMissense <- factor(crispr_data$Color_DeleteriousMissense)
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Other", "cyan3", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)

@@

# Wilcoxon tests: Point mutations

@@

@@

## Grouped by gene  {.tabset .tabset-fade .tabset-pills}

### Deleterious vs. Other

```{r, eval=FALSE}
crispr_signif_del <- compare_means(Score ~ Mutation_Status_Deleterious, group.by = c("Hugo_Symbol"), data = crispr_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_del <- adj_signif(crispr_signif_del)
crispr_signif_del <- crispr_signif_del[order(crispr_signif_del$p),]
saveRDS(crispr_signif_del, "./data_munging/rds/crispr_signif_ptmuts_deleterious_gene.rds")

# write.table(crispr_signif_del, file = "~/Desktop/crispr_signif_ptmuts_deleterious_gene.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_del <- readRDS("./data_munging/rds/crispr_signif_ptmuts_deleterious_gene.rds")

knitr::kable(filter(crispr_signif_del, p < 0.01)[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing deleterious mutant vs other cell lines, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_gene_plot <- ggplot(data = crispr_signif_del) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_gene_plot
```

### Non-silent mutant vs. Other

@@

### Deleterious and missense mutant vs. Other

```{r, eval=FALSE}
crispr_signif_delmis <- compare_means(Score ~ Mutation_Status_DeleteriousMissense, group.by = c("Hugo_Symbol"), data = crispr_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_delmis <- adj_signif(crispr_signif_delmis)
crispr_signif_delmis <- crispr_signif_delmis[order(crispr_signif_delmis$p),]
saveRDS(crispr_signif_delmis, "./data_munging/rds/crispr_signif_ptmuts_deleteriousmissense_gene.rds")

# write.table(crispr_signif_delmis, file = "~/Desktop/crispr_signif_ptmuts_deleteriousmissense_gene.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_delmis <- readRDS("./data_munging/rds/crispr_signif_ptmuts_deleteriousmissense_gene.rds")

knitr::kable(filter(crispr_signif_delmis, p < 0.01)[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing deleterious and missense mutant vs other cell lines, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_gene_delmis_plot <- ggplot(data = crispr_signif_delmis) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_gene_delmis_plot
```

## Grouped by gene and lineage {.tabset .tabset-fade .tabset-pills}

### Deleterious vs. Other

```{r, eval=FALSE}
crispr_signif_del_lineage <- compare_means(Score ~ Mutation_Status_Deleterious, group.by = c("Hugo_Symbol", "group_general_lineage_name"), data = crispr_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_del_lineage <- adj_signif(crispr_signif_del_lineage)
crispr_signif_del_lineage <- crispr_signif_del_lineage[order(crispr_signif_del_lineage$p),]
saveRDS(crispr_signif_del_lineage, "./data_munging/rds/crispr_signif_ptmuts_deleterious_lineage.rds")

# write.table(crispr_signif_del_lineage, file = "~/Desktop/crispr_signif_ptmuts_deleterious_lineage.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_del_lineage <- readRDS("./data_munging/rds/crispr_signif_ptmuts_deleterious_lineage.rds")

knitr::kable(filter(crispr_signif_del_lineage, p < 0.01)[, c("Hugo_Symbol", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing deleterious mutant vs other cell lines by lineage, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_lineage_plot <- ggplot(data = crispr_signif_del_lineage) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_lineage_plot
```

### Non-silent mutant vs. Other

@@

### Deleterious and missense mutant vs. Other

```{r, eval=FALSE}
crispr_signif_delmis_lineage <- compare_means(Score ~ Mutation_Status_DeleteriousMissense, group.by = c("Hugo_Symbol", "group_general_lineage_name"), data = crispr_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_delmis_lineage <- adj_signif(crispr_signif_delmis_lineage)
crispr_signif_delmis_lineage <- crispr_signif_delmis_lineage[order(crispr_signif_delmis_lineage$p),]
saveRDS(crispr_signif_delmis_lineage, "./data_munging/rds/crispr_signif_ptmuts_deleteriousmissense_lineage.rds")

# write.table(crispr_signif_delmis_lineage, file = "~/Desktop/crispr_signif_ptmuts_deleteriousmissense_lineage.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_delmis_lineage <- readRDS("./data_munging/rds/crispr_signif_ptmuts_deleteriousmissense_lineage.rds")

knitr::kable(filter(crispr_signif_delmis_lineage, p < 0.01)[, c("Hugo_Symbol", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing deleterious and missense mutant vs other cell lines by lineage, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_gene_delmis_lineage_plot <- ggplot(data = crispr_signif_delmis_lineage) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_gene_delmis_lineage_plot
```

# Wilcoxon tests: All mutations

## Grouped by gene  {.tabset .tabset-fade .tabset-pills}

### Deleterious vs. Other

```{r, eval=FALSE}
crispr_signif_allmut_del <- compare_means(Score ~ Mutation_Status_Deleterious, group.by = c("Hugo_Symbol"), data = crispr_data, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_allmut_del <- adj_signif(crispr_signif_allmut_del)
crispr_signif_allmut_del <- crispr_signif_allmut_del[order(crispr_signif_allmut_del$p),]
saveRDS(crispr_signif_allmut_del, "./data_munging/rds/crispr_signif_allmut_deleterious_gene.rds")

# write.table(crispr_signif_allmut_del, file = "~/Desktop/crispr_signif_allmut_deleterious_gene.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_allmut_del <- readRDS("./data_munging/rds/crispr_signif_allmut_deleterious_gene.rds")

knitr::kable(filter(crispr_signif_allmut_del, p < 0.01)[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing deleterious mutant vs other cell lines, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_gene_allmut_plot <- ggplot(data = crispr_signif_allmut_del) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_gene_allmut_plot
# ggsave(filename = "./plots_18Q3/crispr/pvals_genes.pdf", plot = wilcox_gene_plot, width = 8, height = 4, device = "pdf")
```

### Non-silent mutant vs. Other

```{r, eval=FALSE}
crispr_signif_allmut_nonsilent <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol"), data = crispr_data, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_allmut_nonsilent <- adj_signif(crispr_signif_allmut_nonsilent)
crispr_signif_allmut_nonsilent <- crispr_signif_allmut_nonsilent[order(crispr_signif_allmut_nonsilent$p),]
saveRDS(crispr_signif_allmut_nonsilent, "./data_munging/rds/crispr_signif_allmut_nonsilent_gene.rds")

# write.table(crispr_signif_allmut_nonsilent, file = "~/Desktop/crispr_signif_allmut_nonsilent_gene.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_allmut_nonsilent <- readRDS("./data_munging/rds/crispr_signif_allmut_nonsilent_gene.rds")

knitr::kable(filter(crispr_signif_allmut_nonsilent, p < 0.01)[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing non-silent mutant vs other cell lines, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_gene_allmut_nonsilent_plot <- ggplot(data = crispr_signif_allmut_nonsilent) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_gene_allmut_nonsilent_plot
```

### Deleterious and missense mutant vs. Other

```{r, eval=FALSE}
crispr_signif_allmut_delmis <- compare_means(Score ~ Mutation_Status_DeleteriousMissense, group.by = c("Hugo_Symbol"), data = crispr_data, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_allmut_delmis <- adj_signif(crispr_signif_allmut_delmis)
crispr_signif_allmut_delmis <- crispr_signif_allmut_delmis[order(crispr_signif_allmut_delmis$p),]
saveRDS(crispr_signif_allmut_delmis, "./data_munging/rds/crispr_signif_allmut_deleteriousmissense_gene.rds")

# write.table(crispr_signif_allmut_delmis, file = "~/Desktop/crispr_signif_allmut_deleteriousmissense_gene.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_allmut_delmis <- readRDS("./data_munging/rds/crispr_signif_allmut_deleteriousmissense_gene.rds")

knitr::kable(filter(crispr_signif_allmut_delmis, p < 0.01)[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing deleterious and missense mutant vs other cell lines, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_gene_allmut_delmis_plot <- ggplot(data = crispr_signif_allmut_delmis) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_gene_allmut_delmis_plot
```

## Grouped by gene and lineage {.tabset .tabset-fade .tabset-pills}

### Deleterious vs. Other

```{r, eval=FALSE}
crispr_signif_allmut_del_lineage <- compare_means(Score ~ Mutation_Status_Deleterious, group.by = c("Hugo_Symbol", "group_general_lineage_name"), data = crispr_data, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_allmut_del_lineage <- adj_signif(crispr_signif_allmut_del_lineage)
crispr_signif_allmut_del_lineage <- crispr_signif_allmut_del_lineage[order(crispr_signif_allmut_del_lineage$p),]
saveRDS(crispr_signif_allmut_del_lineage, "./data_munging/rds/crispr_signif_allmut_deleterious_lineage.rds")

# write.table(crispr_signif_allmut_del_lineage, file = "~/Desktop/crispr_signif_allmut_deleterious_lineage.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_allmut_del_lineage <- readRDS("./data_munging/rds/crispr_signif_allmut_deleterious_lineage.rds")

knitr::kable(filter(crispr_signif_allmut_del_lineage, p < 0.01)[, c("Hugo_Symbol", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing deleterious mutant vs other cell lines by lineage, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_lineage_allmut_plot <- ggplot(data = crispr_signif_allmut_del_lineage) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_lineage_allmut_plot
```

### Non-silent mutant vs. Other

```{r, eval=FALSE}
crispr_signif_allmut_nonsilent_lineage <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol", "group_general_lineage_name"), data = crispr_data, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_allmut_nonsilent_lineage <- adj_signif(crispr_signif_allmut_nonsilent_lineage)
crispr_signif_allmut_nonsilent_lineage <- crispr_signif_allmut_nonsilent_lineage[order(crispr_signif_allmut_nonsilent_lineage$p),]
saveRDS(crispr_signif_allmut_nonsilent_lineage, "./data_munging/rds/crispr_signif_allmut_nonsilent_lineage.rds")

# write.table(crispr_signif_allmut_nonsilent_lineage, file = "~/Desktop/crispr_signif_allmut_nonsilent_lineage.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_allmut_nonsilent_lineage <- readRDS("./data_munging/rds/crispr_signif_allmut_nonsilent_lineage.rds")

knitr::kable(filter(crispr_signif_allmut_nonsilent_lineage, p < 0.01)[, c("Hugo_Symbol", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing non-silent mutant vs other cell lines by lineage, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_lineage_allmut_nonsilent_plot <- ggplot(data = crispr_signif_allmut_nonsilent_lineage) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_lineage_allmut_nonsilent_plot
```

### Deleterious and missense mutant vs. Other

```{r, eval=FALSE}
crispr_signif_allmut_delmis_lineage <- compare_means(Score ~ Mutation_Status_DeleteriousMissense, group.by = c("Hugo_Symbol", "group_general_lineage_name"), data = crispr_data, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_allmut_delmis_lineage <- adj_signif(crispr_signif_allmut_delmis_lineage)
crispr_signif_allmut_delmis_lineage <- crispr_signif_allmut_delmis_lineage[order(crispr_signif_allmut_delmis_lineage$p),]
saveRDS(crispr_signif_allmut_delmis_lineage, "./data_munging/rds/crispr_signif_allmut_deleteriousmissens_lineage.rds")

# write.table(crispr_signif_allmut_delmis_lineage, file = "~/Desktop/crispr_signif_allmut_deleteriousmissense_lineage.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=FALSE}
crispr_signif_allmut_delmis_lineage <- readRDS("./data_munging/rds/crispr_signif_allmut_deleteriousmissens_lineage.rds")

knitr::kable(filter(crispr_signif_allmut_delmis_lineage, p < 0.01)[, c("Hugo_Symbol", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing deleterious vs missense mutant and other cell lines, p < 0.01 (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=4, fig.width=8, eval=FALSE}
wilcox_lineage_allmut_delmis_plot <- ggplot(data = crispr_signif_allmut_delmis_lineage) +
  geom_histogram(aes(x = p, fill = "chartreuse4"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  geom_histogram(aes(x = p.adj, fill = "darkslategray3"), breaks = seq(0, 1, by = 0.025), color = "black", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), labels = seq(0, 1, by = 0.05)) +
  scale_fill_manual(name = "P-values", values = c("chartreuse4" = "chartreuse4", "darkslategray3" = "darkslategray3"), labels = c("Unadjusted", "BH-adjusted")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = c(0.1, 0.85)) +
  labs(x = "BH-adjusted p-values", y = "Frequency")
wilcox_lineage_allmut_delmis_plot
```

@@

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset.tsv", sep = "\t", header = TRUE)
genes_lvlA <- unique(g2p$Gene)
```

>>>>

@@

## Genomic annotations

### CCLE MAF file

@@

# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Other", "Mutant")
maf_df <- unique(subset(maf_df, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Mutation_Status_Nonsilent", "Reference_Allele_Length", "Tumor_Seq_Allele1_Length")))

@@

### CCLE copy number file

@@

### CCLE gene expression file

@@

## CRISPR

### DepMap dependency probabilities

@@

### DepMap cell line metadata

### Merge CRISPR data and annotations

@@

```{r fig.height=12, fig.width=12, eval=FALSE}
crispr_pca_data <- merge(ccl_converter[, c("Broad_ID", "group_general_lineage_name")], crispr, by = "Broad_ID")
crispr_lin <- crispr_pca_data[, 2]
crispr_pca <- prcomp(data.matrix(crispr_pca_data[, 3:ncol(crispr_pca_data)]), scale. = TRUE, center = TRUE)
plot(crispr_pca, type = "l")
biplot(crispr_pca, obs.scale = 1, var.scale = 1, groups = crispr_lin, ellipse = TRUE, circle = TRUE)
ggord(crispr_pca, crispr_pca_data$group_general_lineage_name)
```

@@

# Merge mutation annotations
crispr_muts <- merge(crispr_melt, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Other", Mutation_Status_Nonsilent))

# Summarize number of mutant and Other cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Nonsilent_Other = sum(Mutation_Status_Nonsilent == "Other"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))

@@

# Add Color columns
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Other", "thistle", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)

@@

```{r, eval=FALSE, include=FALSE}
# Code for testing colors on plots
test <- filter(crispr_data, Hugo_Symbol == "KRAS")
ggplot(data = test, aes(x = Mutation_Status_Nonsilent, y = Score, fill = Mutation_Status_Nonsilent)) +
  geom_boxplot() +
  geom_jitter(aes(color = Mutation_Status_Nonsilent)) +
  theme_light() +
  scale_color_manual(values = c("turquoise4", "paleturquoise3")) +
  scale_fill_manual(values = c("turquoise4", "paleturquoise3"))
```

## Drug screens

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)

# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC_CCLE")

# CTRP
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.vars = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC_CTRP")

# GDSC
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC_GDSC")

# Merge all data
drug_data <- merge(merge(ccle_melt, ctrp_melt, by = c("accession_id", "CID", "Drug"), all = TRUE), gdsc_melt, by = c("accession_id", "CID", "Drug"), all = TRUE)
drug_data <- merge(drug_data, ccl_converter, by = "accession_id", all.x = TRUE)
drug_data <- merge(drug_data, maf_df, by = c("CCLE_Name", "Broad_ID"))
drug_data$Hugo_Symbol <- factor(drug_muts$Hugo_Symbol)

saveRDS(drug_data, "./../crispr_lineages_giant_files/drug_data_18Q3.rds", compress = "xz")
```

# Wilcoxon tests

@@

## CRISPR

@@

@@

### Grouped by gene and lineage: Non-silent mutant vs. Other

@@

@@

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_v2_10_10_2018.tsv", sep = "\t", header = TRUE)
genes_lvlA <- unique(g2p$Gene)
```
